---
title: "Lab 6"
author: "Eva Wilson"
date: "11/1/2021"
output: 
  html_document: 
    theme: readable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(broom)
library(equatiomatic)
```

## Example of a rank-based test 

We'll make our own samples, using a pseudorandom generator. 

**Explanation:**

- When you tell R to create a sample of integer values, there is a ***huge*** sequence of values that you can start at (e.g. 3, 4, 30, 2, 1, 15, 6, 8) or pull from. 

- Setting the seed allows us to say, if you are pulling from this sequence to get a sample, then I want you to start at 1414. 

- Note: this is just a "random" sample set. 

- Int stands for integers in this range (from 0-20) pick 15 values, replace means when you take the sample then replace it. 

```{r}
set.seed(1414)
gp_1 <- sample.int(20, size = 15, replace = TRUE)

set.seed(1424)
gp_2 <- sample.int(30, size = 15, replace = TRUE)
```

'Hist' is a function that shortcuts ggplot
```{r}
hist(gp_1)

hist(gp_2)
```
Try a t-test
```{r}
t.test(gp_1, gp_2)
```

If these samples were drawn from populations with the same mean, the probability of taking two random samples with means at least as different as the sample means we found by random chance (taking into account spread, n) is 19.8%. 

Retain (fail to reject) the null hypothesis. >> There is no significant difference in means between group 1 and group 2. 

Warning: people get weirdly upset if you say "accept" the null. 

Now, let's compare this outcome to a rank-based test. (rank based means these data are independent from each other)

## Mann Whitney U unpaired rank-based test 

(HOW DID YOU GET THIS TITLE TO BE A DIFFERENT COLOR FONT)

```{r}
mwu <- wilcox.test(gp_1, gp_2)

mwu
```

Null: These populations have the same RANKS (as opposed to means)

If these samples were drawn from populations with the same ranks of values (medians), the probability of taking two random samples with ranks at least as different as those in our samples is 28%. 

There is no significant difference in ranks (often you'll see medians) between group 1 and group 2 (statistical summary).

Median scores for group 1 (M = 14) and group 2 (M = 12) did not differ significantly. (Mann Whitney U test: U(df) = 86, p = 0.28). 

***Note M = median*** 

These together - a t-test and a mann whitney u test - talk about mean and median, respectively. They can make a strong argument. Just make sure you use consistent language if you only choose one versus the other (e.g. only talk about means if you are going to use t-test, only talk about median if you are going to use Mann Whitney u test). In other words, stay consistent when you talk about central tendency. 

If you don't know what the outcomes are from an object that is a list:
?wilcox.test - scroll to value - this will tell you what the pieces of this list are 

mwu#p.value will show you the actual value of that 

## Linear regression 

Simple linear regression (single dependent variable, a single independent variable)

```{r, include = TRUE}
# include = FALSE means that none of this code will show up in your knitted document
# Make an exploratory plot of penguin body mass versus flipper length (x-axis)
ggplot(data = penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point() +
  geom_smooth(method = "lm") # This is okay for exploring only, but do not just assume a linear relationship makes sense unless you have tested it
```

Find a linear regression model using ordinary least squares describing the relationship between flipper length and body mass for these penguins. 

3 pieces:

- What type of model?
- What is the relationship to model (DV ~ IV(s))
- Where is the data that's used to create this model? 

```{r}
penguin_lm <- lm(body_mass_g ~ flipper_length_mm, data = penguins)
# Make sure you tell it what data you are strating from
```

y = mx +b
y= B1x + B0 + E
y = body mass g 
x = flipper length mm
Body Mass (g) = 49.7(flipper_length) - 5780.80

Remember, penguin_lm$coefficients tells you to find all the values associated with object. Here, we are looking for coefficient associated with flipper length. 
penguin_lm$coefficients[2] = pulls out just the flipper_length_mm from the coefficients 

## broom package returns model outputs as tidy data frames 

```{r}
penguin_lm_tidy <- broom::tidy(penguin_lm)

penguin_lm_tidy

broom::glance(penguin_lm) 


# This gives you measures of the terms of coefficients, the actual estimates, the standard erros, test statitic value. You want this because you might want to put these terms in a summary table of key statistics 

# Glance gets you model wide indicators of fit and uncertainty. 
```

How can I actually include my model equation in a report? 

```{r}
extract_eq(model = penguin_lm, use_coefs = TRUE)

```

```{r}
plot(penguin_lm)
```
Residiual versus fitted - how far these points exist from what the model would predict. Do these look somewhat evenly spread around the predicted line? Overall, these residuals are somewhat evenly dispersed around predictor line.

QQPlot: residuals are normally distributed 

Residuals versus Leverage: what points have a disproportionate weight on the other pointsin a way that might be concerning. 

**All four graphs plot residuals, which is how you determine whether a linear model is appropriate!**
